### 2020-01-29 CalNotes.tex
Bayesian

Let $\epl > 0$ and let us consider the one-dimensional multiscale stochastic differential equation (SDE)
\begin{equation}\label{eq:SDE_MS}
	\d X_t^\epl = -\alpha \cdot V'(X_t^\epl) \dd t - \frac1\epl p'\left(\frac{X_t^\epl}\epl\right) + \sqrt{2\sigma} \dd W_t,
\end{equation}
where $\alpha_i \in \R$, $i = 1, \ldots, N$ and $\sigma > 0$ are the drift and diffusion coefficients respectively and $W_t$ is a standard one-dimensional Brownian motion. The functions $V\colon \R \to \R^N$ and $p\colon \R \to \R$ are slow and fast potentials driving the dynamics of the solution $X_t^\epl$. Given $N \in \N_{>0}$, we assume the slow potential to be of the form
\begin{equation}\label{eq:Potential}
	V(x) = \begin{pmatrix} V_1(x) & V_2(x) & \cdots & V_N(x) \end{pmatrix}^\top,
\end{equation}
for smooth functions $V_i\colon \R \to \R$, $i = 1, \ldots, N$. Moreover, we assume $p$ to be smooth and periodic of period $L$. Theory of homogenization \cite{BLP78} guarantees the existence of an SDE of the form
\begin{equation}\label{eq:SDE_HOM}
	\d X_t^0 = - A \cdot V'(X_t^0) \dd t + \sqrt{2\Sigma} \dd W_t,
\end{equation}
where $W_t$ is the same Brownian motion and where the fast dynamics have been eliminated, such that $X_t^\epl \to X_t^0$ for $\epl\to 0$ in law as random variables in $\mathcal C^0((0, T), \R)$. In particular, we have $A_i = K\alpha_i$ and $\Sigma = K \sigma$, where the coefficient $K$ is given by the formula
\begin{equation}\label{eq:K_HOM}
	K = \int_0^L (1 + \Phi'(y))^2 \, \mu(\d y),
\end{equation}
with 
\begin{equation}
	\mu(\d y) = Z^{-1} e^{-p(y)/\sigma} \dd y, \quad\text{where}\quad Z = \int_0^L e^{-p(y)/\sigma} \dd y,
\end{equation}
and where the function $\Phi$ is the solution of the elliptic partial differential equation
\begin{equation}
	-p'(y)\Phi'(y) + \sigma \Phi''(y) = p'(y), \quad 0 \leq y \leq L,
\end{equation}
endowed with periodic boundary conditions.

\subsection{Failure without pre-processing \corr{rewrite}}

@@

\subsection{Outline}

\section{Posterior formal computations}

The log-posterior density is then given by
\begin{equation}
\begin{aligned}
	\log p(A \mid X) = &-\log Z -\frac1{2\Sigma}\int_0^T A \cdot V'(X_t) \dd X_t - \frac1{4\Sigma} \int_0^T \left(A \cdot V'(X_t) \right)^2 \dd t \\
	&- \frac12 \norm{C_0^{-1/2}(A - A_0)}_2^2.
\end{aligned}
\end{equation}
Since the log-posterior density is quadratic in $A$, the posterior is Gaussian, and it is therefore sufficient to determine its mean and covariance to fully characterize it. We denote by $m_T$ and $C_T$ the mean and covariance matrix, respectively. Let us consider the matrix-valued function $M \colon \R \to \R^{N\times N}$ whose entries are given by
\begin{equation}
	M_{ij} = \frac1{2\Sigma T} \int_0^T V_i'(X_t) \, V_j'(X_t) \dd t, \quad i, j = 1, \ldots, N,
\end{equation}
and the vector-valued function $h \colon \R \to \R^N$ defined by
\begin{equation}
	h_i = \frac1{2\Sigma T} \int_0^T V_i'(X_t) \dd X_t, \quad i = 1, \ldots, N.
\end{equation}
Let us remark that we can rewrite $M$ and $h$ in vector notation as
\begin{equation}
		M = \frac1{2\Sigma T} \int_0^T V'(X_t) \otimes V'(X_t) \dd t, \quad h = \frac1{2\Sigma T} \int_0^T V'(X_t) \dd X_t,
\end{equation}
where $\otimes$ denotes the tensor product in $\R^N$. Therefore, we can rewrite the log-posterior density as
\begin{equation}
\begin{aligned}
	\log p(A \mid X) = &-\log Z - T A \cdot h - \frac{T}{2} A \cdot M A - \frac12 (A - A_0) \cdot C_0^{-1}(A-A_0).
\end{aligned}
\end{equation}
Completing the squares in the log-posterior density, it is possible to show that the posterior is a Gaussian $\mu(A \mid X) = \mathcal N(m_T, C_T)$, whose precision matrix and mean are formally given by
\begin{equation}
\begin{aligned}
	C_T^{-1} &= C_0^{-1} + T M, \\
	C_T^{-1}m_T &= C_0^{-1}A_0 - T h. 
\end{aligned}
\end{equation}
We are now interested in the limit of the posterior distribution for $T \to \infty$. Let us first define the maximum likelihood estimator (MLE) $\widehat A(X)$ of $A$, which is obtained by maximizing the log-likelihood function $\log p(X \mid A)$, and which is hence formally given by
\begin{equation}
	\widehat A(X) = -M^{-1}h.
\end{equation}
We now introduce regularity assumptions on the SDE.
\begin{assumption}\label{as:regularity} The potentials $p$ and $V$ are such that 
	\begin{enumerate}
		\item for all $T > 0$, the symmetric matrix $M$ is positive definite and its minimum eigenvalue satisfies $\lambda_{\min}(M) \geq \bar \lambda > 0$,
		\item 	\corr{other assumptions to be added}
	\end{enumerate}
\end{assumption}

@@

%Let us consider $N = 1$ and write $A_1 = A$ and $V_1 = V$, respectively. The posterior distribution of $A$ given a trajectory $(X_t, 0 \leq t \leq T)$ is in this case a Gaussian $\mathcal N(\bar A_T, \sigma^2_T)$ where
%\begin{equation}
%\begin{aligned}
%	\bar A_T &= \frac{A_0}{1+\sigma_0^2\int_0^T V'(X_t)^2 \dd t} - \frac{\sigma_0^2\int_0^T  V'(X_t) \dd X_t}{1+\sigma_0^2\int_0^T V'(X_t)^2 \dd t}, \\
%	\sigma^{-2}_T &= \sigma_0^{-2} + \int_0^T V'(X_t)^2 \dd t.
%\end{aligned}
%\end{equation}
%We will in the following sections analyse the convergence of the posterior, both in case data from the multiscale process is not treated and in case they are pre-processed.

@@

The estimator of the homogenized drift coefficient $A$ we analyse is given by the formula
\begin{equation}\label{eq:AHatMixed}
	\widehat A^\epl(T) = - \widetilde M^{-1} \widetilde h,
\end{equation}
where
\begin{equation}
	\widetilde M = \frac{1}{T} \int_0^T V'(Z^\epl_t) \otimes V'(X^\epl_t) \dd t \qquad \text{and} \qquad \widetilde h = \frac{1}{T} \int_0^T V'(Z^\epl_t) \dd X^\epl_t.
\end{equation}
In practice, one evaluation of $V'(\cdot)$ is computed with the filter, while the differential and the second evaluation of $V'(\cdot)$ are maintained with the original process $X^\epl$.

\begin{remark} Let us remark that in the one-dimensional case $N = 1$, the estimator reduces to
\begin{equation}
\widehat A^\epl(T) = - \frac{\int_0^T V'(Z^\epl_t) \dd X^\epl_t}{\int_0^T V'(Z^\epl_t) V'(X^\epl_t) \dd t}.
\end{equation}
\end{remark}

\begin{theorem}\label{thm:mainTheorem} Let $\widehat A^\epl(T)$ be defined in \eqref{eq:AHatMixed}. Then, if \corr{Assumptions}, 
	\begin{equation}\label{eq:theoremFirst}
	\lim_{\epl \to 0} \lim_{T \to \infty} \widehat A^\epl(T) = A, \quad \text{a.s.},
	\end{equation}
	where $A$ is the drift coefficient of the homogenized equation \eqref{eq:SDE_MS}.
%   Moreover, the following central limit theorem holds
%	\begin{equation}\label{eq:theoremSecond}
%	\lim_{\epl \to 0} \lim_{T \to \infty} \sqrt{T} \left(A - \widehat A^\epl(T) \right) = \Lambda, \quad \text{in law}, 
%	\end{equation}
%	where $\Lambda \sim \mathcal N(0, \gamma^2)$ and where 
%	\begin{equation}
%	\gamma = \sqrt{2\sigma} \frac{\sqrt{\E^{\rho^0}[V'(Z)^2]}}{\E^{\rho^0}[V'(Z)V'(X)]}.
%	\end{equation}
\end{theorem}

@@

>>>>

Let $\epl > 0$ and let us consider the one-dimensional multiscale stochastic differential equation (SDE)
\begin{equation}\label{eq:SDE_MS}
	\d X_t^\epl = -\alpha \cdot V'(X_t^\epl) \dd t - \frac1\epl p'\left(\frac{X_t^\epl}\epl\right) \dd t + \sqrt{2\sigma} \dd W_t,
\end{equation}
where $\alpha_i \in \R$, $i = 1, \ldots, N$ and $\sigma > 0$ are the drift and diffusion coefficients respectively and $W_t$ is a standard one-dimensional Brownian motion. The functions $V\colon \R \to \R^N$ and $p\colon \R \to \R$ are slow and fast potentials driving the dynamics of the solution $X_t^\epl$. Given $N \in \N_{>0}$, we assume the slow potential to be of the form
\begin{equation}\label{eq:Potential}
	V(x) = \begin{pmatrix} V_1(x) & V_2(x) & \cdots & V_N(x) \end{pmatrix}^\top,
\end{equation}
for smooth functions $V_i\colon \R \to \R$, $i = 1, \ldots, N$. Moreover, we assume $p$ to be smooth and periodic of period $L$. Theory of homogenization \cite{BLP78} guarantees the existence of an SDE of the form
\begin{equation}\label{eq:SDE_HOM}
	\d X_t^0 = - A \cdot V'(X_t^0) \dd t + \sqrt{2\Sigma} \dd W_t,
\end{equation}
where $W_t$ is the same Brownian motion and where the fast dynamics have been eliminated, such that $X_t^\epl \to X_t^0$ for $\epl\to 0$ in law as random variables in $\mathcal C^0((0, T), \R)$. In particular, we have $A_i = K\alpha_i$ and $\Sigma = K \sigma$, where the coefficient $K$ is given by the formula
\begin{equation}\label{eq:K_HOM}
	K = \int_0^L (1 + \Phi'(y))^2 \, \mu(\d y),
\end{equation}
with 
\begin{equation}
	\mu(\d y) = Z^{-1} e^{-p(y)/\sigma} \dd y, \quad\text{where}\quad Z = \int_0^L e^{-p(y)/\sigma} \dd y,
\end{equation}
and where the function $\Phi$ is the solution of the elliptic partial differential equation
\begin{equation}
	-p'(y)\Phi'(y) + \sigma \Phi''(y) = p'(y), \quad 0 \leq y \leq L,
\end{equation}
endowed with periodic boundary conditions.

\section{The Bayesian setting}\label{sec:Bayesian}

The log-posterior density is then given by
\begin{equation}
\begin{aligned}
	\log p(A \mid X) = &-\log Z -\frac1{2\Sigma}\int_0^T A \cdot V'(X_t) \dd X_t - \frac1{4\Sigma} \int_0^T \left(A \cdot V'(X_t) \right)^2 \dd t \\
	&- \frac12 \norm{C_0^{-1/2}(A - A_0)}_2^2.
\end{aligned}
\end{equation}
Since the log-posterior density is quadratic in $A$, the posterior is Gaussian, and it is therefore sufficient to determine its mean and covariance to fully characterize it. We denote by $m_T$ and $C_T$ the mean and covariance matrix, respectively. Let us consider the matrix-valued function $M \colon \R \to \R^{N\times N}$ whose entries are given by
\begin{equation}
	M_{ij} = \frac1{2\Sigma T} \int_0^T V_i'(X_t) \, V_j'(X_t) \dd t, \quad i, j = 1, \ldots, N,
\end{equation}
and the vector-valued function $h \colon \R \to \R^N$ defined by
\begin{equation}
	h_i = \frac1{2\Sigma T} \int_0^T V_i'(X_t) \dd X_t, \quad i = 1, \ldots, N.
\end{equation}
Let us remark that we can rewrite $M$ and $h$ in vector notation as
\begin{equation}
		M = \frac1{2\Sigma T} \int_0^T V'(X_t) \otimes V'(X_t) \dd t, \quad h = \frac1{2\Sigma T} \int_0^T V'(X_t) \dd X_t,
\end{equation}
where $\otimes$ denotes the tensor product in $\R^N$. Therefore, we can rewrite the log-posterior density as
\begin{equation}
	\log p(A \mid X) = -\log Z - T A \cdot h - \frac{T}{2} A \cdot M A - \frac12 (A - A_0) \cdot C_0^{-1}(A-A_0).
\end{equation}
Completing the squares in the log-posterior density, it is possible to show that the posterior is a Gaussian $\mu(A \mid X) = \mathcal N(m_T, C_T)$, whose precision matrix and mean are formally given by
\begin{equation}
\begin{aligned}
	C_T^{-1} &= C_0^{-1} + T M, \\
	C_T^{-1}m_T &= C_0^{-1}A_0 - T h. 
\end{aligned}
\end{equation}
We are now interested in the limit of the posterior distribution for $T \to \infty$. Let us first define the maximum likelihood estimator (MLE) $\widehat A(X)$ of $A$, which is obtained by maximizing the log-likelihood function $\log p(X \mid A)$, and which is hence formally given by
\begin{equation}
	\widehat A(X) = -M^{-1}h.
\end{equation}
We now introduce regularity assumptions on the SDE.
\begin{assumption}\label{as:regularity} The potentials $p$ and $V$ are such that 
	\begin{enumerate}
		\item for all $T > 0$, the symmetric matrix $M$ is positive definite and its minimum eigenvalue satisfies $\lambda_{\min}(M) \geq \bar \lambda > 0$,
		\item 	\corr{other assumptions to be added}
	\end{enumerate}
\end{assumption}

\section{Failure without pre-processing}

@@

@@

@@

The estimator of the homogenized drift coefficient $A$ we analyse is given by the formula
\begin{equation}\label{eq:AHatMixed}
	\widehat A^\epl(T) = - \widetilde M^{-1} \widetilde h,
\end{equation}
where
\begin{equation}\label{eq:MandHTilde}
	\widetilde M = \frac{1}{T} \int_0^T V'(Z^\epl_t) \otimes V'(X^\epl_t) \dd t \qquad \text{and} \qquad \widetilde h = \frac{1}{T} \int_0^T V'(Z^\epl_t) \dd X^\epl_t.
\end{equation}
In practice, one evaluation of $V'(\cdot)$ is computed with the filter, while the differential and the second evaluation of $V'(\cdot)$ are maintained with the original process $X^\epl$.

\begin{remark} Let us remark that in the one-dimensional case $N = 1$, the estimator can be written as
\begin{equation}
\widehat A^\epl(T) = - \frac{\int_0^T V'(Z^\epl_t) \dd X^\epl_t}{\int_0^T V'(Z^\epl_t) V'(X^\epl_t) \dd t}.
\end{equation}
\end{remark}

In the one-dimensional case, it is clear that the maximum likelihood estimator based on $X^\epl$ alone, i.e., the quantity
\begin{equation}
	\widehat A^\epl(T) = - \frac{\int_0^T V'(X^\epl_t) \dd X^\epl_t}{\int_0^T V'(X^\epl_t)^2 \dd t}.
\end{equation}
is well defined for $V(x) \neq 0$. Equivalently, for a $N$-dimensional parameter, a sufficient assumption for the well-posedness of the estimator is to require the symmetric matrix $M$ defined as
\begin{equation}
	M = \frac{1}{T} \int_0^T V'(X^\epl_t) \otimes V'(X^\epl_t) \dd t,
\end{equation}
to be positive definite. In the following lemma, we

\begin{lemma} The matrix $\widetilde M$ defined in \eqref{eq:MandHTilde} is invertible almost surely. \end{lemma}
\begin{proof}

\begin{theorem}\label{thm:mainTheorem} Let $\widehat A^\epl(T)$ be defined in \eqref{eq:AHatMixed}. Then
	\begin{equation}\label{eq:theoremFirst}
	\lim_{\epl \to 0} \lim_{T \to \infty} \widehat A^\epl(T) = A, \quad \text{a.s.},
	\end{equation}
	where $A$ is the drift coefficient of the homogenized equation \eqref{eq:SDE_MS}.
%   Moreover, the following central limit theorem holds
%	\begin{equation}\label{eq:theoremSecond}
%	\lim_{\epl \to 0} \lim_{T \to \infty} \sqrt{T} \left(A - \widehat A^\epl(T) \right) = \Lambda, \quad \text{in law}, 
%	\end{equation}
%	where $\Lambda \sim \mathcal N(0, \gamma^2)$ and where 
%	\begin{equation}
%	\gamma = \sqrt{2\sigma} \frac{\sqrt{\E^{\rho^0}[V'(Z)^2]}}{\E^{\rho^0}[V'(Z)V'(X)]}.
%	\end{equation}
\end{theorem}

@@

In this section we show how the filtering approach can be employed in a Bayesian framework. As in Section \ref{sec:Bayesian} let us consider a Gaussian prior $\mu_0 = \mathcal N(A_0, C_0)$. Then, the density $p^\epl(A \mid X)$ of the posterior distribution $\mu^\epl(A \mid X^\epl)$ given by multiscale data satisfies
\begin{equation}
	p^\epl(A \mid X) = \frac1{Z^\epl} \, p^\epl(X \mid A) \, p_0(A),
\end{equation}
where $Z^\epl$ is the normalization term and where the likelihood function is given by
\begin{equation}
	p^\epl(X \mid A) = \exp\left(-\frac{I^\epl(X\mid A)}{2\Sigma} \right), 
\end{equation}
with
\begin{equation}
	I^\epl(X\mid A) = \int_0^T A \cdot V'(X_t^\epl) \dd X_t^\epl + \frac12 \int_0^T \left( A \cdot V'(X_t^\epl) \right)^2 \dd t.
\end{equation}
Let us consider the likelihood function $\widetilde p^\epl(X \mid A)$ modified to include the filtered data $Z^\epl$ given by
\begin{equation}
	\widetilde p^\epl(X \mid A) = \exp\left(-\frac{\widetilde I^\epl(X\mid A)}{2\Sigma} \right), 
\end{equation}
with
\begin{equation}
	\widetilde I^\epl(X\mid A) = \int_0^T A \cdot V'(Z_t^\epl) \dd X_t^\epl + \frac12 \int_0^T \left( A \cdot V'(X_t^\epl) \right) \left( A \cdot V'(Z_t^\epl) \right) \dd t.
\end{equation}
Moreover, let us consider the modified posterior distribution $\widetilde \mu^\epl(A \mid X)$ with density
\begin{equation}
	\widetilde p^\epl(A \mid X) = \frac1{\widetilde Z^\epl} \, \widetilde p^\epl(X \mid A) \, p_0(A),
\end{equation}
Let us remark that employing \eqref{eq:MandHTilde} we can write 
\begin{equation}
	\log \widetilde p^\epl(A \mid X) = -\log \widetilde Z^\epl - T A \cdot \widetilde h - \frac{T}{2} A \cdot \widetilde M A - \frac12 (A - A_0) \cdot C_0^{-1}(A-A_0).
\end{equation}
With respect to Section \ref{sec:Bayesian}, the matrix $\widetilde M$ is not symmetric. Nonetheless, we remark that
\begin{equation}
	A \cdot \widetilde M A = A \cdot \widetilde M_S A,
\end{equation}
where $\widetilde M_S$ is the symmetric part of $\widetilde M$, i.e.,
\begin{equation}
	\widetilde M_S = \frac12 \left(\widetilde M + \widetilde M^\top\right).
\end{equation}
Hence, we can rewrite 
\begin{equation}	
	\log \widetilde p^\epl(A \mid X) = -\log \widetilde Z^\epl - T A \cdot \widetilde h - \frac{T}{2} A \cdot \widetilde M_S A - \frac12 (A - A_0) \cdot C_0^{-1}(A-A_0).
\end{equation}
This shows that the posterior $\widetilde \mu^\epl$ is a Gaussian $\mathcal N(\widetilde m_T, \widetilde C_T)$ where the covariance and mean can be obtained as 
\begin{equation}
\begin{aligned}
	\widetilde C_T^{-1} &= C_0^{-1} + T \widetilde M, \\
	\widetilde C_T^{-1} \widetilde m_T &= C_0^{-1}A_0 - T \widetilde h. 
\end{aligned}
\end{equation}

In the following theorem we show that the posterior we obtain 
\begin{theorem} Under
\end{theorem}
